Typing a linear \picalc

### 2020-04-11 paper/paper.tex
Rewrite the intro to leftover typing

% Do not use italics for definitions
\theoremstyle{definition}
\newtheorem{nidefinition}[theorem]{Definition}

@@

\todo{stress what leftover typing is as a technique}
\todo{syntax directed typing rules}

In a resource-aware calculi, typing rules modify the context of multiplicities under which their premises are typed -- e.g. in $\send{x}{y}P$ the multiplicity context under which $P$ is typed needs to be modified at at $x$ and $y$.
This context transformations can be modeled as functions on contexts, however threading these transformations through the proofs on typing rules is cumbersome, and these transformations necessitate of a myriad of additional lemmas.
Additionally, if the typing rules do not record the result of their actions on contexts, managing parallel composition entails a context split for which the user is required to provide an existential proof at the cutting point \todo{rephrase}.

An alternative is to keep track in the typing judgment of the way in which the context has been transformed.
One way of doing so is by mentioning both the \emph{input contexts} and the \emph{output contexts} over which the processes are typed.
Parallel composition then uses the output context of the left hand process as the input context of the right hand.
As an additional benefit, the associativity of parallel composition becomes trivial to prove.

@@

\todo{op but for contexts}
\begin{theorem}
  Let $\opctx{\Gamma_l}{\Delta}{\Xi_l}$ and $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
  If $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$ then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
\end{theorem}

@@

@@

@@

@@

@@

>>>>

% Do not use italics in definitions and theorems
\theoremstyle{definition}
\newtheorem{nidefinition}[theorem]{Definition}
\newtheorem{nitheorem}[theorem]{Theorem}

@@

We use leftover typing \cite{} to encode our typing rules.
The insight of this approach is to use input and output contexts in the typing judgment.
Our typing judgment on processes are of the form $\types{\gamma}{\Gamma}{P}{\Delta}$ (introduced in Figure \ref{fig:types}).
This assertion is in fact both a typing judgment on $\gamma$ and a usage judgment on $\Gamma$ and $\Delta$.
The typing context $\gamma$ needs to be explicit because the usage context $\Gamma$ and $\Delta$ do not in any way depend on it -- in fact, they are only related by their common length.
Using both an input and an output usage context has several advantages.
For one, the usage modifications effected by both input and output can be enclosed within a self-containing typing and usage judgment for variable references -- also with an output context.
But more crucially, we avoid requesting an explicit constructive proof of context splitting for parallel composition:
if $P$ is typed under $\Gamma_l$ and $Q$ under $\Gamma_r$, then to type $\comp{P}{Q}$ under $\Gamma_m$ we would need a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$.
The search for such a proof can be automated, but its existence is still required.
Moreover, providing a typing derivation for $P$ gets the user no usage annotations to type $Q$.
By using output usage contexts, we get the leftovers from the derivation on $P$ and use them to type $Q$.

\todo{syntax directed typing rules}

@@

\todo{op but for contexts}
\begin{nitheorem}
  Let $\opctx{\Gamma_l}{\Delta}{\Xi_l}$ and $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
  If $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$ then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
\end{nitheorem}

@@

\begin{nitheorem}
  If $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
\end{nitheorem}

@@

\begin{nitheorem}
  If $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\Unused \; i \; P$, then $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
\end{nitheorem}

\begin{remark}
  The proof that index $i$ is unused in $P$ must be syntactical and cannot rely on the usage annotation of index $i$ being preserved between $\Gamma$ and $\Xi$: process $P$ can send variable reference $i$ with multiplicity $\zero$ over a channel.
\end{remark}

@@

\begin{nitheorem}
  If $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
\end{nitheorem}

@@

\begin{nitheorem}
  If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
\end{nitheorem}

@@

\subsection{Subject Reduction}

\begin{nitheorem}
  If $P \reduce{c} Q$ and $\types{\gamma}{\Gamma'}{P}{\Xi}$ where $\opctx{\Gamma'}{\ell_{\#} @ c}{\Gamma}$,
  then $\types{\gamma}{\Gamma}{Q}{\Xi}$
\end{nitheorem}
